---
title: "Reproducible Research Course Project 1"
author: "Alex Fennell"
date: "1/7/2022"
output: html_document
---

```{r libraries, message=FALSE}
    library(dplyr)
    library(ggplot2)
```

## Loading and preprocessing the data

```{r download data}
fileurl<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if (!file.exists("CP1data.zip")){
    download.file(fileurl,"CP1data.zip",method="curl")
    unzip("CP1data.zip")
}
act<-read.csv("activity.csv")
act$date<-as.Date(act$date, format="%Y-%m-%d")
str(act)

```
With the data in the desired format now, it is necessary to get a quick idea of
the data using summary
```{r}
summary(act)
```
This shows a good amount of missing data ~13% so to calculate the mean it is 
necessary to remove the NA values
## What is mean total number of steps taken per day?
```{r}
totalsteps<-act %>%
    group_by(date) %>%
    na.omit(steps) %>%
    summarise(totsteps=sum(steps))
    qplot(totalsteps$totsteps,
          main="Total Steps Taken Each Day",
          xlab="Total Steps")+
    theme(plot.title=element_text(hjust=.5))
```

```{r mean steps}
    mean(act$steps,na.rm=TRUE)
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
